---
title: "2016 US Presidential Elections"
date: "`r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=8)
library(plyr)  # `plyr` must be loaded before `dplyr`
library(dplyr)
library(lubridate)
```

#### Days until Election Day: `r (as.Date("2016-11-08") - Sys.Date()) %>% as.numeric`

<br>

The data in this report (unless otherwise noted) comes from using the `pollstR` package.  It allows for direct access to the *HuffPost Pollster* API using `R`.  The code appears on [Github](https://github.com/daniel-park/Election2016).

<br>


## National Polling Estimate  
The current estimate on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`.  
```{r US_table}
setwd("~/Dropbox/DataScience/Projects/Election2016")

library(pollstR)

# Chart name for various national polls.
slug <- "2016-general-election-trump-vs-clinton-vs-johnson"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

## National Polling Data {.tabset .tabset-fade .tabset-pills}

### National Polls  

National support for the two main US presidential candidates according to various polling organizations.
```{r US_polls}
source("Election2016_Polls.R")

polls.plotly
```

### National Estimates  

The line represents *HuffPost Pollster's* estimate for overall support based on polling data.

```{r US_est}
# Comes from `Election2016_Polls.R` script.
est.plotly 
```

```{r eval=FALSE, echo=FALSE}
# Fully annotated time series

# marker1b will denote election day
# 1b is different from 1a in the text placement
marker1b <- list(x="2016-11-15",
                y=midpoint,
                text="Election Day <br> November 08",
                font=list(size=10),
                showarrow=FALSE)

# Trump convention speech
marker2 <- list(x=as.Date("2016-07-21"), 
                y=filter(est.daily, date==as.Date("2016-07-21") & choice=="Trump")$value,
                text="Convention <br> Speech",
                arrowsize=0.7,
                ay=45,
                ax=10,
                font=list(size=10),
                showarrow=TRUE)

# Clinton convention speech
marker3 <- list(x=as.Date("2016-07-28"), 
                y=filter(est.daily, date==as.Date("2016-07-28") & choice=="Clinton")$value,
                text="Convention <br> Speech",
                arrowsize=0.7,
                ay=-30,
                ax=-10,
                font=list(size=10),
                showarrow=TRUE)               



plot.labels <- list(marker1b, marker2, marker3)

# Annotated polling estimate plot
plot_ly(data=est.daily.top2,
            x=date,
            y=value,
            mode="lines + markers + text",
            color=choice,
            colors=c("red", "blue"),
        showlegend=FALSE) %>%
  layout(title="National Polling Overall Estimates", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", 
                    #ticktext=ticktext, 
                    #tickvals=tickvals, 
                    #tickmode="array",
                    tickangle=-45, 
                    rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```

<br>

## Electoral Map  

This is the current electoral map according to estimates by *HuffPost Pollster*.  Estimates are made for states with sufficient polling data.

```{r ElectoralMap}
source("Election2016_ElectoralMap.R")

election.map
```

<br>

## Campaign Stops {.tabset .tabset-fade .tabset-pills}

The records begin on July 31, 2016 (100 days until Election Day).  The information is being collected daily from [here](https://www.donaldjtrump.com/schedule) and [here](https://www.willhillarywin.com/hillary-clintons-events/).


### Today

```{r tour_today}
source("Election2016_ScheduleScrape.R")

source("Election2016_TourMap.R")

map.today

schedule.state <- "US"
schedule.date.begin <- Sys.Date()
schedule.date.end <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```




### Yesterday

```{r tour_yesterday}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.yesterday

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 1
schedule.date.end   <- Sys.Date() - 1

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last Three Days

```{r tour_last3}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last3days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 2
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last 7 Days

```{r tour_last7}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last7days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 6
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last 30 Days  
```{r tour_last30}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last30days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 29
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Since July 31, 2016

```{r tour_all}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.100days

schedule.state <- "US"
schedule.date.begin <- as.Date("2016-07-31")
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

<br>
<br>

## Battleground States  {.tabset .tabset-fade .tabset-pills}

There are roughly eleven swing states.

A [source](http://www.nytimes.com/2016/07/31/us/politics/donald-trump-presidential-race.html) within the Trump campaign identified four must-win states: Ohio, Pennsylvania, Florida, and North Carolina.

<br>

### Florida

**Present Polling Estimate for Florida on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r FL_table}
slug <- "2016-florida-presidential-general-election-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>




#### Florida Polling Data {.tabset .tabset-fade .tabset-pills}

##### Florida Polls

```{r FL_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Florida Estimates
```{r FL_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to Florida  

```{r FL_trips}
schedule.state <- "FL"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### North Carolina  

**Present Polling Estimate for North Carolina on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NC_table}
slug <- "2016-north-carolina-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### North Carolina Polling Data {.tabset .tabset-fade .tabset-pills}

##### North Carolina Polls

```{r NC_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### North Carolina Estimates
```{r NC_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to North Carolina  

##### Democrats
```{r NC_trips}
schedule.state <- "NC"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Ohio  

**Present Polling Estimate for Ohio on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r OH_table}
slug <- "2016-ohio-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### Ohio Polling Data {.tabset .tabset-fade .tabset-pills}

##### Ohio Polls

```{r OH_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Ohio Estimates
```{r OH_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to Ohio  

```{r OH_trips}
party.name <- "Democratic"
schedule.state <- "OH"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Pennsylvania  

While races are typically close in Pennsylvania, Democrats have won this state in the past six presidential elections.  The last time that Republicans won this state was in 1988.

**Present Polling Estimate for Pennsylvania on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r PA_table}
slug <- "2016-pennsylvania-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### Pennsylvania Polling Data {.tabset .tabset-fade .tabset-pills}


##### Pennsylvania Polls

```{r PA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Pennsylvania Estimates
```{r PA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly

```

<br>

#### Campaign Visits to Pennsylvania {.tabset .tabset-fade .tabset-pills}  

```{r PA_trips}
schedule.state <- "PA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Other {.tabset .tabset-fade .tabset-pills}

#### Colorado

**Present Polling Estimate for Colorado on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r CO_table}
slug <- "2016-colorado-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Colorado Polling Data {.tabset .tabset-fade .tabset-pills}

###### Colorado Polls

```{r CO_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Colorado Estimates
```{r CO_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Colorado    

```{r CO_trips}
schedule.state <- "CO"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Iowa

**Present Polling Estimate for Iowa on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r IA_table}
slug <- "2016-iowa-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Iowa Polling Data {.tabset .tabset-fade .tabset-pills}

###### Iowa Polls

```{r IA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Iowa Estimates
```{r IA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Iowa  

```{r IA_trips}
schedule.state <- "IA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Michigan

**Present Polling Estimate for Michigan on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r MI_table}
slug <- "2016-michigan-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Michigan Polling Data {.tabset .tabset-fade .tabset-pills}

###### Michigan Polls

```{r MI_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Michigan Estimates
```{r MI_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Michigan  

```{r MI_trips}
schedule.state <- "MI"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Nevada

**Present Polling Estimate for Nevada on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NV_table}
slug <- "2016-nevada-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Nevada Polling Data {.tabset .tabset-fade .tabset-pills}

###### Nevada Polls

```{r NV_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Nevada Estimates
```{r NV_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Nevada  

```{r NV_trips}
schedule.state <- "NV"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### New Hampshire

**Present Polling Estimate for New Hampshire on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NH_table}
slug <- "2016-new-hampshire-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### New Hampshire Polling Data {.tabset .tabset-fade .tabset-pills}

###### New Hampshire Polls

```{r NH_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### New Hampshire Estimates
```{r NH_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to New Hampshire 

```{r NH_trips}
schedule.state <- "NH"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Virginia

**Present Polling Estimate for Virginia on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r VA_table}
slug <- "2016-virginia-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Virginia Polling Data {.tabset .tabset-fade .tabset-pills}

###### Virginia Polls

```{r VA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Virginia Estimates
```{r VA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Virginia 

```{r VA_trips}
schedule.state <- "VA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Wisconsin

**Present Polling Estimate for Wisconsin on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r WI_table}
slug <- "2016-wisconsin-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Wisconsin Polling Data {.tabset .tabset-fade .tabset-pills}

###### Wisconsin Polls

```{r WI_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Wisconsin Estimates
```{r WI_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>
 
##### Campaign Visits to Wisconsin

```{r WI_trips}
schedule.state <- "WI"

source("Election2016_ScheduleTable.R")

schedule.DT
```


 <br>
 <br>