---
title: "2016 US Presidential Elections"
date: "`r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=8)
library(plyr)  # `plyr` must be loaded before `dplyr`
library(dplyr)
library(lubridate)
```


<br>

The data in this report (unless otherwise noted) comes from using the `pollstR` package.  It allows for direct access to the *HuffPost Pollster* API using `R`.  The code for this report appears on [Github](https://github.com/daniel-park/Election2016).

<br>

#### Days until Election Day: `r (as.Date("2016-11-08") - Sys.Date()) %>% as.numeric`

<br>

## National Polling Estimate  
The current estimate on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`.  
```{r US_table}
setwd("~/Dropbox/DataScience/Projects/Election2016")

library(pollstR)

# Chart name for various national polls.
slug <- "2016-general-election-trump-vs-clinton-vs-johnson"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

## National Polling Data {.tabset .tabset-fade .tabset-pills}

### National Polls  

National support for the two main US presidential candidates according to various polling organizations.
```{r US_polls}
source("Election2016_Polls.R")

polls.plotly
```

### National Estimates  

The line represents *HuffPost Pollster's* estimate for overall support based on polling data.

```{r US_est}
# Comes from `Election2016_Polls.R` script.
est.plotly 
```

<br>

## Electoral Map  

This is the current electoral map according to estimates by *HuffPost Pollster*.  Estimates are made for states with sufficient polling data.

```{r ElectoralMap}
source("Election2016_ElectoralMap.R")

election.map
```

<br>

## Campaign Stops {.tabset .tabset-fade .tabset-pills}

The records begin on July 31, 2016 (100 days until Election Day).  The information is being collected daily from [here](https://www.donaldjtrump.com/schedule) and [here](https://www.willhillarywin.com/hillary-clintons-events/).

### Today

```{r tour_today}
source("Election2016_ScheduleScrape.R")

source("Election2016_TourMap.R")

map.today

schedule.state <- "US"
schedule.date.begin <- Sys.Date()
schedule.date.end <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```




### Yesterday

```{r tour_yesterday}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.yesterday

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 1
schedule.date.end   <- Sys.Date() - 1

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last Three Days

```{r tour_last3}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last3days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 2
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last 7 Days

```{r tour_last7}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last7days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 6
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Last 30 Days  
```{r tour_last30}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.last30days

schedule.state <- "US"
schedule.date.begin <- Sys.Date() - 29
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```

### Since July 31, 2016

```{r tour_all}
# See scripts
#   `Election2016_ScheduleScrape.R`
#   `Election2016_TourMap.R`
#   for source.
map.100days

schedule.state <- "US"
schedule.date.begin <- as.Date("2016-07-31")
schedule.date.end   <- Sys.Date()

source("Election2016_ScheduleTable.R")

schedule.DT
```


## Campaign Focus
```{r focus}

library(plotly)
library(plyr)
library(dplyr)

setwd("~/Dropbox/DataScience/Projects/Election2016")

# Create file name which contains yesterday's date
file.name <- paste0("Election2016_csv/Election2016_TourSchedule", 
                    format(Sys.Date(), "%b%d"), 
                    ".csv")

# Load file whose name contains yesterday's date
# Note that it is located in a different file than the current working 
#   directory
schedule.df <- read.csv(file=file.name, stringsAsFactors=FALSE)

pie.party.df <- schedule.df %>% group_by(state, party) %>% summarise(count = n())

state.abb[51] <- "DC"
state.name[51] <- "Washington D.C."

# Change from state abbreviation to full name
pie.party.df$name <- sapply(pie.party.df$state, FUN = function(x) {
  state.name[which(x == state.abb)] 
}) %>% unlist

pie.dem.df <- filter(pie.party.df, party == "Democratic") %>% 
  arrange(desc(count)) %>% 
  as.data.frame %>% 
  mutate(hover = paste0(name, " (", count, ")"))
  
pie.party.df <- arrange(pie.party.df, party, desc(count)) %>% 
  as.data.frame %>% 
  mutate(hover = paste0(name, " (", count, ")"))

pie.dem.df <- pie.party.df %>% filter(party == "Democratic")
pie.repub.df <- pie.party.df %>% filter(party == "Republican")

pie.dem.df <- rbind(pie.dem.df %>% slice(1:7) %>% 
                      mutate(hover = paste(name, "<br>", paste(count, "events"))),
                    data.frame(state = "other",
            party = "Democratic",
            count = sum(pie.dem.df$count[8:nrow(pie.dem.df)]),
            name = "other",
            hover = paste(pie.dem.df$hover[8:nrow(pie.dem.df)], collapse = "<br>")
          )
)
  
pie.repub.df <- rbind(pie.repub.df %>% slice(1:7) %>% 
                      mutate(hover = paste(name, "<br>", paste(count, "events"))),
                    data.frame(state = "other",
            party = "Republican",
            count = sum(pie.repub.df$count[8:nrow(pie.repub.df)]),
            name = "other",
            hover = paste(pie.repub.df$hover[8:nrow(pie.repub.df)], collapse = "<br>")
          )
)









plot_ly(data = pie.dem.df, 
        labels=state, 
        values=count, 
        text = hover,
        textinfo = "label", # labeling on chart
        hoverinfo = "text+percent",
        type="pie", 
        hole=0.5,
        domain = list(x = c(0, 0.4), y = c(0.15, 1)),
        marker = list(colors = c('rgb(15,95,160)', 'rgb(33,113,181)', 'rgb(66,146,198)', 'rgb(107,174,214)', 'rgb(158,202,225)', 'rgb(198,219,239)', 'rgb(222,235,247)', 'rgb(235,245,252)')),
        showlegend = FALSE) %>% 
  add_trace(data = pie.repub.df, 
        labels=state, 
        values=count, 
        text = hover,
        textinfo = "label", # labeling on chart
        hoverinfo = "text+percent",
        type="pie", 
        hole=0.5,
        domain = list(x = c(0.55, 0.95), y = c(0.15, 1)),
        marker = list(colors = c('rgb(165,15,21)', 'rgb(203,24,29)', 'rgb(239,59,44)', 'rgb(251,106,74)', 'rgb(252,146,114)', 'rgb(252,187,161)', 'rgb(254,224,210)','rgb(255,245,240)')),  
        showlegend = FALSE) %>% 
  layout(title = "Campaign Focus by Party", 
         annotations = list(
            list(x = 0.13 , y = .99, showarrow = F, text = "Democrats", font = list(size = 18)),
            list(x = 0.84 , y = .99, showarrow = F, text = "Republicans", font = list(size = 18)),
            list(x = 0.16 , y = .17, showarrow = F, text = paste("Total:", sum(pie.dem.df$count)), font = list(size = 14)),
            list(x = 0.81 , y = .17, showarrow = F, text = paste("Total:", sum(pie.repub.df$count)), font = list(size = 14))
            )
  )


```
<br>
<br>
<br>
<br>

## Battleground States  {.tabset .tabset-fade .tabset-pills}

There are roughly eleven swing states.

A [source](http://www.nytimes.com/2016/07/31/us/politics/donald-trump-presidential-race.html) within the Trump campaign identified four must-win states: Ohio, Pennsylvania, Florida, and North Carolina.

<br>

### Florida

**Present Polling Estimate for Florida on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r FL_table}
slug <- "2016-florida-presidential-general-election-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>




#### Florida Polling Data {.tabset .tabset-fade .tabset-pills}

##### Florida Polls

```{r FL_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Florida Estimates
```{r FL_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to Florida  

```{r FL_trips}
schedule.state <- "FL"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### North Carolina  

**Present Polling Estimate for North Carolina on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NC_table}
slug <- "2016-north-carolina-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### North Carolina Polling Data {.tabset .tabset-fade .tabset-pills}

##### North Carolina Polls

```{r NC_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### North Carolina Estimates
```{r NC_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to North Carolina  

```{r NC_trips}
schedule.state <- "NC"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Ohio  

**Present Polling Estimate for Ohio on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r OH_table}
slug <- "2016-ohio-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### Ohio Polling Data {.tabset .tabset-fade .tabset-pills}

##### Ohio Polls

```{r OH_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Ohio Estimates
```{r OH_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

#### Campaign Visits to Ohio  

```{r OH_trips}
schedule.state <- "OH"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Pennsylvania  

While races are typically close in Pennsylvania, Democrats have won this state in the past six presidential elections.  The Republicans' last victory here was in 1988.

**Present Polling Estimate for Pennsylvania on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r PA_table}
slug <- "2016-pennsylvania-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

#### Pennsylvania Polling Data {.tabset .tabset-fade .tabset-pills}


##### Pennsylvania Polls

```{r PA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

##### Pennsylvania Estimates
```{r PA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly

```

<br>

#### Campaign Visits to Pennsylvania {.tabset .tabset-fade .tabset-pills}  

```{r PA_trips}
schedule.state <- "PA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


### Other {.tabset .tabset-fade .tabset-pills}

#### Colorado

**Present Polling Estimate for Colorado on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r CO_table}
slug <- "2016-colorado-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Colorado Polling Data {.tabset .tabset-fade .tabset-pills}

###### Colorado Polls

```{r CO_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Colorado Estimates
```{r CO_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Colorado    

```{r CO_trips}
schedule.state <- "CO"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Iowa

**Present Polling Estimate for Iowa on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r IA_table}
slug <- "2016-iowa-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Iowa Polling Data {.tabset .tabset-fade .tabset-pills}

###### Iowa Polls

```{r IA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Iowa Estimates
```{r IA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Iowa  

```{r IA_trips}
schedule.state <- "IA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Michigan

**Present Polling Estimate for Michigan on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r MI_table}
slug <- "2016-michigan-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Michigan Polling Data {.tabset .tabset-fade .tabset-pills}

###### Michigan Polls

```{r MI_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Michigan Estimates
```{r MI_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Michigan  

```{r MI_trips}
schedule.state <- "MI"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Nevada

**Present Polling Estimate for Nevada on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NV_table}
slug <- "2016-nevada-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Nevada Polling Data {.tabset .tabset-fade .tabset-pills}

###### Nevada Polls

```{r NV_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Nevada Estimates
```{r NV_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Nevada  

```{r NV_trips}
schedule.state <- "NV"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### New Hampshire

**Present Polling Estimate for New Hampshire on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r NH_table}
slug <- "2016-new-hampshire-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### New Hampshire Polling Data {.tabset .tabset-fade .tabset-pills}

###### New Hampshire Polls

```{r NH_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### New Hampshire Estimates
```{r NH_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to New Hampshire 

```{r NH_trips}
schedule.state <- "NH"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Virginia

**Present Polling Estimate for Virginia on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r VA_table}
slug <- "2016-virginia-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Virginia Polling Data {.tabset .tabset-fade .tabset-pills}

###### Virginia Polls

```{r VA_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Virginia Estimates
```{r VA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>

##### Campaign Visits to Virginia 

```{r VA_trips}
schedule.state <- "VA"

source("Election2016_ScheduleTable.R")

schedule.DT
```


#### Wisconsin

**Present Polling Estimate for Wisconsin on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`**
```{r WI_table}
slug <- "2016-wisconsin-president-trump-vs-clinton"

source("Election2016_EstTable.R")

candidate.DT
```

<br>

##### Wisconsin Polling Data {.tabset .tabset-fade .tabset-pills}

###### Wisconsin Polls

```{r WI_polls}
source("Election2016_Polls.R")

# Polling data
polls.plotly
```

###### Wisconsin Estimates
```{r WI_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
est.plotly
```

<br>
 
##### Campaign Visits to Wisconsin

```{r WI_trips}
schedule.state <- "WI"

source("Election2016_ScheduleTable.R")

schedule.DT
```


 <br>
 <br>