---
title: "Election Data"
date: "`r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=8)
setwd("~/Dropbox/DataScience/Projects/Election2016")
```

The `pollstR` package.


```{r init}
library(pollstR)

# Available charts, list of 2
#   List 1: `$charts` shows available charts
#   List 2: `$estimates` shows today's estimates for candidates, 
#           approval ratings, etc.
charts <- pollster_charts(page=1) # is this necessary?
# Result:
# View(charts$charts)  # listing by name


```

###National Polling Average on `r paste0(format(Sys.Date(), format='%B %d'), ', 2016')`
```{r US_table}
library(pollstR)
# Chart name for various national polls.
slug <- "2016-general-election-trump-vs-clinton-vs-johnson"



source("Election2016_EstTable.R")

# Cannot source `datatable()` from an R script.
DT::datatable(data=est.today,
              options=list(dom="t"),
              caption = htmltools::tags$caption(
                style = 'caption-side: bottom; text-align: left;',
                htmltools::em("Source: HuffPost Pollster"))
)
```

## Polls {.tabset}

### National Polls

```{r US_polls}
source("Election2016_Polls.R")

library(plotly)
# Polling data
plot_ly(data=all.polls.long, 
        x=med_date, 
        y=value, 
        mode="markers", 
        marker=list(color=choice_color),
        hoverinfo="text",
        text=hover_info,
        color=choice) %>% 
  #colors=c("red", "blue")
  layout(title="National Polls", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
```

### National Estimates
```{r US_est}
source("Election2016_Est.R")


# Overlay estimates with polling data
plot_ly(data=est.daily.top2,
        x=date,
        y=value,
        #line=list(color=choice_color),
        mode="lines + markers + text",
        color=choice,
        colors=c("red", "blue"),
        showlegend=FALSE) %>%
  add_trace(data=all.polls.long, 
            x=med_date, 
            y=value, 
            marker=list(color=choice_color),
            mode="markers", 
            hoverinfo="none",
            color=choice 
            #colors=c("red", "blue")
  ) %>%
  layout(title="National Polls and Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
               

```

```{r eval=FALSE, echo=FALSE}
# Fully annotated time series

# marker1b will denote election day
# 1b is different from 1a in the text placement
marker1b <- list(x="2016-11-15",
                y=midpoint,
                text="Election Day <br> November 11",
                font=list(size=10),
                showarrow=FALSE)

# Trump convention speech
marker2 <- list(x=as.Date("2016-07-21"), 
                y=filter(est.daily, date==as.Date("2016-07-21") & choice=="Trump")$value,
                text="Convention <br> Speech",
                arrowsize=0.7,
                ay=45,
                ax=10,
                font=list(size=10),
                showarrow=TRUE)

# Clinton convention speech
marker3 <- list(x=as.Date("2016-07-28"), 
                y=filter(est.daily, date==as.Date("2016-07-28") & choice=="Clinton")$value,
                text="Convention <br> Speech",
                arrowsize=0.7,
                ay=-30,
                ax=-10,
                font=list(size=10),
                showarrow=TRUE)               



plot.labels <- list(marker1b, marker2, marker3)

# Annotated polling estimate plot
plot_ly(data=est.daily.top2,
            x=date,
            y=value,
            mode="lines + markers + text",
            color=choice,
            colors=c("red", "blue"),
        showlegend=FALSE) %>%
  layout(title="National Polling Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", 
                    #ticktext=ticktext, 
                    #tickvals=tickvals, 
                    #tickmode="array",
                    tickangle=-45, 
                    rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```

## Electoral Map
```{r ElectoralMap, eval=FALSE}
#data(state)



map.df <- data.frame(state=state.abb, 
                     name=state.name, 
                     clinton_value=numeric(50),
                     trump_value=numeric(50),
                     leader=character(50), 
                     lead=numeric(50),
                     stringsAsFactors=FALSE)


# slugs for states that have polling data
state.slug.df <- filter(charts$charts, topic=="2016-president") %>%
  filter(state!="US") %>%
  select(slug, state)

map.df <- merge(map.df, state.slug.df, by="state", all=TRUE)

map.df$slug <- ifelse(test=is.na(map.df$slug), 
                      yes="",
                      no=map.df$slug)


# for every row in `map.df`
for (i in 1:nrow(map.df)) {
  
  # if there exists estimates for that state
  if (any(map.df$slug[i]==state.slug.df$slug, na.rm=TRUE)) {
    
    tryCatch({
      
    # extract the data for that state
    state.est.today <- pollster_chart(map.df$slug[i])$estimates # today's estimates
    
    # take today's estimated voter support for each candidate
    # record in `map.df`
    map.df$clinton_value[i] <- filter(state.est.today, choice=="Clinton")$value
    map.df$trump_value[i] <- filter(state.est.today, choice=="Trump")$value
    
    # take the candidate leading the state (always first entry)
    # record candidate's name in `map.df`
    map.df$leader[i] <- state.est.today$choice[1]
    
    # take difference between voter support for both candidates
    # record difference in `map.df`
    map.df$lead[i]   <- state.est.today$value[1] - state.est.today$value[2]
    }, error=function(e) {
      
      #cat("\n", "Error related to slug", map.df$slug[i],
      #    ":",conditionMessage(e), "\n")
      }
    )
  }
}



# MAP

# add column for hover labels on map
map.df <- mutate(map.df, hover=paste(state, "<br>", "Clinton", clinton_value, "<br>",
                                      "Trump", trump_value))

# Need to superimpose two maps, so separate data
clinton.map.df <- filter(map.df, leader=="Clinton")
trump.map.df <- filter(map.df, leader=="Trump")

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)


election.map <- plot_ly(data=clinton.map.df, 
                        z = lead, 
                        text = hover, 
                        hoverinfo="text",
                        locations=state,
                        type = 'choropleth',
                        locationmode = 'USA-states', 
                        color = lead, 
                        colors = 'Blues', 
                        showscale=FALSE) %>%
  add_trace(data=trump.map.df, 
            z = lead,
            text=hover,
            hoverinfo="text",
            locations=state,
            type='choropleth',
            locationmode='USA-states',
            color=lead, 
            colors='Reds', 
            showscale=FALSE) %>%
  layout(title="Electoral Map",
         geo=g)

election.map

```




## Campaign Stops {.tabset}

The following maps track campaign stops. The data begins on July 31, 2016 - 100 days until the Election Day.  

### By Candidate

```{r eval=FALSE}
# Schedule MAP
source("Election2016_ScheduleScrape.R")

library(leaflet)


repub.icon <- makeIcon(iconUrl="https://upload.wikimedia.org/wikipedia/commons/9/9b/Republicanlogo.svg",
                       iconWidth = 25, iconHeight = 25)
dem.icon <- makeIcon(iconUrl="http://www.standupamericaus.org/sua/wp-content/uploads/2012/05/Democrat-Logo-psd355111.png", 
                     iconWidth = 25, iconHeight = 25)

leaflet() %>%
  addTiles() %>%
  addMarkers(data=filter(schedule.df, clinton==TRUE), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             group="Clinton (D)",
             icon=dem.icon) %>%
  addMarkers(data=filter(schedule.df, kaine==TRUE), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             group="Kaine (D)",
             icon=dem.icon) %>%
  addMarkers(data=filter(schedule.df, trump==TRUE), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             group="Trump (R)",
             icon=repub.icon) %>%  
  addMarkers(data=filter(schedule.df, pence==TRUE), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             group="Pence (R)",
             icon=repub.icon) %>%
  addLayersControl(overlayGroups=c("Clinton (D)", "Kaine (D)", 
                                   "Trump (R)", "Pence (R)"),
                   options = layersControlOptions(collapsed = FALSE)) 


```

### By Day

```{r eval=FALSE}
# Partitioning campaign stops by day, week, and all
# Some days, no candidate makes a campaign stop
# In that case, feeding plotly an empty data set causes an error
# Therefore, you must create if-else conditions to prevent this from happening.

# `map.not.today` includes all stops except for today.
map.not.today <- leaflet() %>%
  addTiles() %>% 
  addMarkers(data=filter(schedule.df, party=="republican"), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             icon=repub.icon,
             group="All") %>% 
  addMarkers(data=filter(schedule.df, party=="democrat"), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             icon=dem.icon,
             group="All") %>%
  addMarkers(data=filter(schedule.df, 
                         date <= Sys.Date() & 
                           date > Sys.Date() - 7  & 
                           party=="republican"), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             icon=repub.icon,
             group="Last 7 Days") %>%
  addMarkers(data=filter(schedule.df, 
                         date <= Sys.Date() & 
                           date > Sys.Date() - 7  & 
                           party=="democrat"), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             icon=dem.icon,
             group="Last 7 Days") %>%
  addLayersControl(baseGroups=c("Today", "Last 7 Days", "All"), 
                   options = layersControlOptions(collapsed = FALSE))




if (filter(schedule.df, date == Sys.Date() & party=="republican") %>% nrow > 0
    &
    filter(schedule.df, date == Sys.Date() & party=="democrat") %>% nrow > 0
    ) {
  map.not.today %>%
    addMarkers(data=filter(schedule.df, date==Sys.Date() & party=="republican"),
               lng=~longitude, lat=~latitude,
               popup=~popup_label,
               icon=repub.icon,
               group="Today") %>%
    addMarkers(data=filter(schedule.df, date==Sys.Date() & party=="democrat"), 
             lng=~longitude, lat=~latitude,
             popup=~popup_label,
             icon=dem.icon,
             group="Today")
  } else if (filter(schedule.df, date == Sys.Date() & 
                    party=="republican") %>%
    nrow > 0) {
      map.not.today %>%
    addMarkers(data=filter(schedule.df, date==Sys.Date() & party=="republican"),
               lng=~longitude, lat=~latitude,
               popup=~popup_label,
               icon=repub.icon,
               group="Today")
    } else if (filter(schedule.df, date == Sys.Date() & 
           party=="democrat") %>%
    nrow > 0) {
      map.not.today %>%
        addMarkers(data=filter(schedule.df, date==Sys.Date() & party=="democrat"),
               lng=~longitude, lat=~latitude,
               popup=~popup_label,
               icon=dem.icon,
               group="Today")
    } else {
      map.not.today
    }



```


## Battleground States  {.tabset}
A [source](http://www.nytimes.com/2016/07/31/us/politics/donald-trump-presidential-race.html) within the Trump campaign designated four must-win states: Ohio, Pennsylvania, Florida, and North Carolina.

### Florida
```{r FL_table}
slug <- "2016-florida-presidential-general-election-trump-vs-clinton"

source("Election2016_EstTable.R")

# Cannot source `datatable()` from an R script.
DT::datatable(data=est.today,
              options=list(dom="t"),
              caption = htmltools::tags$caption(
                style = 'caption-side: bottom; text-align: left;',
                htmltools::em("Source: HuffPost Pollster"))
)
```

#### Florida Polling Data {.tabset}

##### Florida Polls

```{r FL_polls}
source("Election2016_Polls.R")

# Polling data
plot_ly(data=all.polls.long, 
        x=med_date, 
        y=value, 
        mode="markers", 
        marker=list(color=choice_color),
        hoverinfo="text",
        text=hover_info,
        color=choice) %>% 
  #colors=c("red", "blue")
  layout(title="Florida Polls", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
```

##### Florida Estimates
```{r FL_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
plot_ly(data=est.daily.top2,
        x=date,
        y=value,
        #line=list(color=choice_color),
        mode="lines + markers + text",
        color=choice,
        colors=c("red", "blue"),
        showlegend=FALSE) %>%
  add_trace(data=all.polls.long, 
            x=med_date, 
            y=value, 
            marker=list(color=choice_color),
            mode="markers", 
            hoverinfo="none",
            color=choice 
            #colors=c("red", "blue")
  ) %>%
  layout(title="Florida Polls and Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```


### North Carolina
```{r NC_table}
slug <- "2016-north-carolina-president-trump-vs-clinton"

source("Election2016_EstTable.R")

# Cannot source `datatable()` from an R script.
DT::datatable(data=est.today,
              options=list(dom="t"),
              caption = htmltools::tags$caption(
                style = 'caption-side: bottom; text-align: left;',
                htmltools::em("Source: HuffPost Pollster"))
)
```

#### North Carolina Polling Data {.tabset}

##### North Carolina Polls

```{r NC_polls}
source("Election2016_Polls.R")

# Polling data
plot_ly(data=all.polls.long, 
        x=med_date, 
        y=value, 
        mode="markers", 
        marker=list(color=choice_color),
        hoverinfo="text",
        text=hover_info,
        color=choice) %>% 
  #colors=c("red", "blue")
  layout(title="North Carolina Polls", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
```

##### North Carolina Estimates
```{r NC_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
plot_ly(data=est.daily.top2,
        x=date,
        y=value,
        #line=list(color=choice_color),
        mode="lines + markers + text",
        color=choice,
        colors=c("red", "blue"),
        showlegend=FALSE) %>%
  add_trace(data=all.polls.long, 
            x=med_date, 
            y=value, 
            marker=list(color=choice_color),
            mode="markers", 
            hoverinfo="none",
            color=choice 
            #colors=c("red", "blue")
  ) %>%
  layout(title="North Carolina Polls and Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```

### Ohio
```{r OH_table}
slug <- "2016-ohio-president-trump-vs-clinton"

source("Election2016_EstTable.R")

# Cannot source `datatable()` from an R script.
DT::datatable(data=est.today,
              options=list(dom="t"),
              caption = htmltools::tags$caption(
                style = 'caption-side: bottom; text-align: left;',
                htmltools::em("Source: HuffPost Pollster"))
)
```

#### Ohio Polling Data {.tabset}

##### Ohio Polls

```{r OH_polls}
source("Election2016_Polls.R")

# Polling data
plot_ly(data=all.polls.long, 
        x=med_date, 
        y=value, 
        mode="markers", 
        marker=list(color=choice_color),
        hoverinfo="text",
        text=hover_info,
        color=choice) %>% 
  #colors=c("red", "blue")
  layout(title="Ohio Polls", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
```

##### Ohio Estimates
```{r OH_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
plot_ly(data=est.daily.top2,
        x=date,
        y=value,
        #line=list(color=choice_color),
        mode="lines + markers + text",
        color=choice,
        colors=c("red", "blue"),
        showlegend=FALSE) %>%
  add_trace(data=all.polls.long, 
            x=med_date, 
            y=value, 
            marker=list(color=choice_color),
            mode="markers", 
            hoverinfo="none",
            color=choice 
            #colors=c("red", "blue")
  ) %>%
  layout(title="Ohio Polls and Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```

### Pennsylvania
```{r PA_table}
slug <- "2016-pennsylvania-president-trump-vs-clinton"

source("Election2016_EstTable.R")

# Cannot source `datatable()` from an R script.
DT::datatable(data=est.today,
              options=list(dom="t"),
              caption = htmltools::tags$caption(
                style = 'caption-side: bottom; text-align: left;',
                htmltools::em("Source: HuffPost Pollster"))
)
```

#### Pennsylvania Polling Data {.tabset}

##### Pennsylvania Polls

```{r PA_polls}
source("Election2016_Polls.R")

# Polling data
plot_ly(data=all.polls.long, 
        x=med_date, 
        y=value, 
        mode="markers", 
        marker=list(color=choice_color),
        hoverinfo="text",
        text=hover_info,
        color=choice) %>% 
  #colors=c("red", "blue")
  layout(title="Pennsylvania Polls", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))
```

##### Pennsylvania Estimates
```{r PA_est}
#source("Election2016_Est.R")
# Overlay estimates with polling data
plot_ly(data=est.daily.top2,
        x=date,
        y=value,
        #line=list(color=choice_color),
        mode="lines + markers + text",
        color=choice,
        colors=c("red", "blue"),
        showlegend=FALSE) %>%
  add_trace(data=all.polls.long, 
            x=med_date, 
            y=value, 
            marker=list(color=choice_color),
            mode="markers", 
            hoverinfo="none",
            color=choice 
            #colors=c("red", "blue")
  ) %>%
  layout(title="Pennsylvania Polls and Average", 
         annotations=plot.labels,
         shapes=list(vert.line),
         xaxis=list(title="Date", tickangle=-45, rangeslider=list(type="date")),
         yaxis=list(title="", ticksuffix="%"))

```